---
title: "LITA vs BITA paper"
output:
  html_document:
    df_print: paged
---

This notebook will contain code for cleaning and then analysis of this dataset.

Going to use the whole dataset to redo the analysis for the LITA - BITA paper.

Steps:

1. Get the data
2. Clean and get the patients that i want
3. Use the earlier code to do both matched and unmatched analysis
4. ROMA-like analysis.

```{r}
library(easypackages)
libraries(c(
  "tidyverse", "haven", "readxl",
  "survival", "lubridate", "WeightIt",
  "cobalt", "rms", "Hmisc", "survminer",
  "cmprsk"
))

library(cobalt)
library(Matching)
library(MatchIt)
library(WeightIt)
library(twang)
library(MatchThem)
library(naniar)

library(surv2sampleComp)

```



```{r}
# get some useful functions...



trim_var <- function(x){
  
  # stop if the vector is not numeric
  
  stopifnot(is.numeric(x))
  
  # now to obtain the 1st and 99th percentile.
  
  a <- quantile(x, 0.01)
  b <- quantile(x, 0.99)
  
  # now to replace smallest values with a and 
  # largest values with b
  
  y <- ifelse(x < a, a, 
         ifelse(x > b, b, x))
  
  return(y)
}
```

get the data:

```{r}

df <- read_csv("D:/lita_bita3/big2.csv")

# now to see the data
# keep only unique rows here.

length(unique(df$Pat_ID))

df2 <- df[!duplicated(df$Pat_ID), ]

# df2 now contains the data / 3756 patients.

# we want to do the whole analysis using all these
# patients separated by sita and bita.

# see how many bita

names(df2) <- tolower(names(df2))

df2 %>% count(bima)

# 2911 sita and 845 bita.
# that is going to be our main subgroup variable.

# for the whole analysis, we will keep all 
# patients.
# we will limit patients only for the ROMA study.

# confirm that no radials.

df2 %>% count(radialis)

summary(df2$pre_lvef)

# graft details:

# 0 - no grafts
# 1 - vein
# 3 - lima 
# 4 - rima 

```

So, for the whole group analysis, will not do any limits. But should limit according to routine statistical guidelines for correct values.

```{r}

# limit some cont variables to meaningful ceilings.

df2$creat <- with(df2, ifelse(creat > 5, 5, creat))

df2$pre_lvef <- with(df2, ifelse(pre_lvef > 60, 60, pre_lvef))

# los days

describe(df2$los_days)

# use the trim_var function.

df2$postop_los <- trim_var(df2$los_days)

df2 %>% count(gender) # convert to female

df2$female <- df2$gender - 1

df2 %>% count(female)

```


Now to continue to do the tables as before.

```{r}
# because we want to compare bita vs lita, we will remove those with single vessel disease and also those that have redo.
# especially because we are going to do propensity matching.

df2 %>% count(cad_disease)

df3 <- df2 %>% filter(cad_disease %in% c(2,3)) # now kept only those with either dvd or tvd.

# to remove also those that are redo.

df3 %>% count(prior_cardiac_surgery)

df3 <- df3 %>% filter(prior_cardiac_surgery == 0)


# now df3 contains lita vs bita , dvd/tvd, no prior cardiac surgery.


# define obese as bmi > 30

df3$obese <- with(df3, ifelse(bmi > 30, 1, 0))

df3 %>% count(obese)

# df3 is the complete dataset prior to propensity matching.

# write_csv(df3, 
#          "D:/lita_bita3/complete_dataset.csv" )
```



Now to make the tables 1, 2, 3 and then do propensity matching.

```{r}

# get df3 here again, so that can do tables 1 & 2.

df <- read_csv("D:/lita_bita3/complete_dataset.csv")

# some small changes before tables...

df$diabetes[df$diabetes == 8]<- 0

df$left_main_disease[df$left_main_disease == 2]<- 0

df <- df %>% filter(distals_n != 1)

# now to do table 1.


```




```{r}

# table 1.

vars <- c(
"age_at_surgery", "gender",  "height_cm", "weight_kg", 
"bmi", "diabetes", "art_hypertension",  
"smoker", "hyperlipemia", "copd", "pad", "creat", "pre_dialysis", 
 "pre_lvef", "pre_lvef_function", "elective", "cad_disease", 
"left_main_disease", "prior_pci", "pre_stroke", "bima")

factors <- c(
 "gender",   "diabetes", "art_hypertension",  
"smoker", "hyperlipemia", "copd", "pad", "pre_dialysis", 
  "pre_lvef_function", "elective", "cad_disease", 
"left_main_disease", "prior_pci", "pre_stroke", "bima")

t1 <- tableone::CreateTableOne(vars = vars,
                               factorVars = factors,
                               data = df,
                               strata = c("bima"))


t1 <- print(t1, nonnormal = c("age","creat","pre_lvef", "bmi"))


# write.csv(t1,
#           "D:/lita_bita3/tables/table1_wc.csv")
```




```{r}

# now for the operative details...

vars <- c("opcab", "conversion_to_onpump", "onpump_beating_heart", 
"no_touch_aorta",  "length_of_surgery", "bypass_time", "cross_clamp_time",  "distals_n", "distals_svg_n", "svg_grafts", "seq_n", "proximals_n", "y_graft_n", "bima")

factors <- c("opcab", "conversion_to_onpump", "onpump_beating_heart", 
"no_touch_aorta", 
 "distals_n", "distals_svg_n", "svg_grafts","seq_n", "proximals_n", "y_graft_n", "bima")


t2 <- tableone::CreateTableOne(vars = vars,
                               factorVars = factors,
                               data = df,
                               strata = c("bima"))


t2 <- print(t2, nonnormal = c("length_of_surgery", "bypass_time", "cross_clamp_time"))

# write.csv(t2, 
#           "D:/lita_bita3/tables/table2_wc.csv")

```


Now to see the postoperative early outcomes for the whole group.

```{r}

vars <- c("post_low_cardiac_output", 
"post_iabp", "post_ecmo", "post_mi", "post_reanimation", "post_cardiac_arrhythmia", 
"post_redo_heart", "post_rethorax_bleeding", "post_stroke", "post_dswi", 
"post_sepsis", "died_inhouse", "postop_los", "bima")

factors <- c("post_low_cardiac_output", 
"post_iabp", "post_ecmo", "post_mi", "post_reanimation", "post_cardiac_arrhythmia", 
"post_redo_heart", "post_rethorax_bleeding", "post_stroke", "post_dswi", 
"post_sepsis", "died_inhouse")

t3 <- tableone::CreateTableOne(vars = vars,
                               factorVars = factors,
                               data = df,
                               strata = c("bima"))


t3 <- print(t3, nonnormal = c("postop_los"))

# write.csv(t3, 
#           "D:/lita_bita3/tables/table3_wc.csv")

```


Now to do the survival unadjusted and then do propensity score methods.


```{r}

df$surv_years <- (df$survival_days + 1)/365.24

s <- survfit(Surv(surv_years, died) ~ bima, data = df)

summary(s, times = c(0,5,10,15))

survdiff(Surv(surv_years, died) ~ bima, data = df)

```


Plot for survival unadjusted and then save the plot.

```{r}
 
 a <- ggsurvplot(s,
     surv.scale = "percent",
     risk.table = T,
     xlim = c(0, 15),
     break.x.by = 5,
     conf.int = T,
     legend.labs = c("SITA", "BITA"),
     palette = "lancet",
     censor.size = 0)
 
 
 ggsave(plot = print(a), 
        filename = "D:/lita_bita3/figures/surv_wc.tiff",
      height = 6,
      width = 7,
      units = "in",
      dpi = 1200,
      device = "tiff")


```


Get more unadjusted survival parameters using surv2sample ...

```{r}

res_wc <- surv2sampleComp::surv2sample(
  time = df$surv_years,
  status = df$died,
  arm = df$bima,
  tau = 15,
  timepoints = c(5,10,15)
)

```


```{r}

# now to see the outcome of these parameters for each group and then
# to see the differences between outcomes.

res_wc$group0

```

Same results for the bima group.

```{r}

res_wc$group1

```

Now to see comparisons...

```{r}

res_wc$contrast.all

```
RMST for unadjusted data..

```{r}

res_wc$RMST

```


Now to do the propensity matching and then again see outcome for early
and late mortality in the matched cohort.

```{r}


formula <- bima ~ age_at_surgery + female + diabetes + obese + art_hypertension + 
  copd + pad + pre_stroke + pre_lvef + creat + left_main_disease + cad_disease + prior_pci + hyperlipemia + pre_dialysis

match1 <- matchit(
  data = df, formula = formula,
  method = "nearest", caliper = 0.25,
  replace = F, ratio = 5)

match1

summary(match1)

bt <- bal.tab(match1, un = T)

vn <- var.names(bt, type = "df", file = "D:/lita_bita3/varnames.csv")

new_vn = read.csv("D:/lita_bita3/varnames2.csv")

lp_whole <- 
  
  love.plot(bt, abs = T,un = T,
          grid = T, thresholds = 0.1, colors = c("tomato4","gray"),
          var.names = new_vn)

# save this plot 


ggsave(lp_whole, filename = "D:/lita_bita3/lp_whole.tiff",
       height = 7, width = 5, units = "in", dpi = 900)



# to get the dataset out and save it.

df_wm <- match.data(match1)

df_wm %>% count(bima)
          
# now to save this dataset.

write_csv(df_wm,
"D:/lita_bita3/matched_data.csv")


```

Now to do the tables for matched dataset.


```{r}

df_wm <- read_csv("D:/lita_bita3/matched_data.csv")


# table 1.


vars <- c(
"age_at_surgery", "gender",  
"bmi", "diabetes", "art_hypertension",  
"smoker", "hyperlipemia", "copd", "pad", "creat", "pre_dialysis", 
 "pre_lvef", "pre_lvef_function", "elective", "cad_disease", 
"left_main_disease", "prior_pci", "pre_stroke", "bima")

factors <- c(
 "gender",   "diabetes", "art_hypertension",  
"smoker", "hyperlipemia", "copd", "pad", "pre_dialysis", 
  "pre_lvef_function", "elective", "cad_disease", 
"left_main_disease", "prior_pci", "pre_stroke", "bima")

t1 <- tableone::CreateTableOne(vars = vars,
                               factorVars = factors,
                               data = df_wm,
                               strata = c("bima"))


t1 <- print(t1, nonnormal = c("age","creat","pre_lvef", "bmi"))


write.csv(t1,
            "D:/lita_bita3/tables/table1_matched_wm.csv")

```

Now for operative details and then postoperative outcomes...

```{r}

vars <- c("opcab", "conversion_to_onpump", "onpump_beating_heart", 
"no_touch_aorta",  "length_of_surgery", "bypass_time", "cross_clamp_time",  "distals_n", "distals_svg_n", "svg_grafts", "seq_n", "proximals_n", "y_graft_n", "bima")

factors <- c("opcab", "conversion_to_onpump", "onpump_beating_heart", 
"no_touch_aorta", 
 "distals_n", "distals_svg_n", "svg_grafts","seq_n", "proximals_n", "y_graft_n", "bima")


t2 <- tableone::CreateTableOne(vars = vars,
                               factorVars = factors,
                               data = df_wm,
                               strata = c("bima"))


t2 <- print(t2, nonnormal = c("length_of_surgery", "bypass_time", "cross_clamp_time"))

write.csv(t2, 
           "D:/lita_bita3/tables/table2_matched_wm.csv")

```


Postoperative outcomes ...

```{r}


vars <- c("post_low_cardiac_output", 
"post_iabp", "post_ecmo", "post_mi", "post_reanimation", "post_cardiac_arrhythmia", 
"post_redo_heart", "post_rethorax_bleeding", "post_stroke", "post_dswi", 
"post_sepsis", "died_inhouse", "postop_los")

factors <- c("post_low_cardiac_output", 
"post_iabp", "post_ecmo", "post_mi", "post_reanimation", "post_cardiac_arrhythmia", 
"post_redo_heart", "post_rethorax_bleeding", "post_stroke", "post_dswi", 
"post_sepsis", "died_inhouse")

t3 <- tableone::CreateTableOne(vars = vars,
                               factorVars = factors,
                               data = df_wm,
                               strata = c("bima"))


t3 <- print(t3, nonnormal = c("postop_los"))

write.csv(t3, 
           "D:/lita_bita3/tables/table3_matched_wm.csv")



```


Graph for the matched data.

```{r}



s_wm <- survfit(Surv(surv_years, died) ~ bima, data = df_wm)

summary(s_wm, times = c(0,5,10,15))

survdiff(Surv(surv_years, died) ~ bima, data = df_wm)


```


Now to see the graph for the adjusted data.

```{r}


 
 a_wm <- ggsurvplot(s_wm,
     surv.scale = "percent",
     risk.table = T,
     xlim = c(0, 15),
     break.x.by = 5,
     conf.int = T,
     legend.labs = c("SITA", "BITA"),
     palette = "lancet",
     censor.size = 0)
 
 
 ggsave(plot = print(a_wm), 
        filename = "D:/lita_bita3/figures/surv_matched_wm.tiff",
      height = 6,
      width = 7,
      units = "in",
      dpi = 1200,
      device = "tiff")




```

